---
title: "p8105_hw6_xl3054"
author: "Xinhui Lin (xl3054)"
date: "2025-12-03"
output: github_document
---

# Problem 1
```{r}
library(tidyverse)
homicide = read_csv("data/homicide-data.csv") %>% 
  janitor::clean_names() %>%
  mutate(city_state = paste0(city, ", ", state),
         solved_bin = ifelse(disposition == "Closed by arrest", 1, 0),
         victim_age = ifelse(victim_age %in% c("Unknown", ""), NA, victim_age),
         victim_age = as.numeric(victim_age),
         victim_sex  = factor(victim_sex, levels = c("Female", "Male")),
         victim_race = factor(victim_race)) %>% 
  filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")),
         victim_race %in% c("White", "Black"))

## for Baltimore
homicide %>% 
  filter(city_state == "Baltimore, MD") %>% 
  glm(solved_bin ~ victim_age + victim_sex + victim_race, 
      data = . , family = binomial) %>% 
  broom::tidy(conf.int = TRUE, exponentiate = TRUE) %>% 
  filter(term == "victim_sexMale") %>% 
  select(term, estimate, conf.low, conf.high) %>% 
  knitr::kable(digits = 4)

## for each city
homicide_or = homicide %>% 
  group_by(city_state) %>% 
  nest() %>% 
  mutate(model = purrr::map(data, 
                            ~ glm(solved_bin ~ victim_age + victim_sex + victim_race,
                                  data   = .x,
                                  family = binomial)),
         tidied = purrr::map(model, ~ broom::tidy(.x, conf.int = TRUE, exponentiate = TRUE))) %>% 
  unnest(tidied) %>% 
  filter(term == "victim_sexMale") %>% 
  select(city_state, term, estimate, conf.low,conf.high) %>% 
  arrange(desc(estimate))

homicide_or %>% 
  knitr::kable(digits = 4)

homicide_or %>% 
  ggplot(aes(x = reorder(city_state, estimate), y = estimate)) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +
  coord_flip() +
  labs(x = "City", y = "Adjusted odds ratio (male vs female)",
       title = "Adjusted OR for solving homicides (Male vs Female)") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 7))
```

This plot shows the adjusted odds ratios for solving homicides involving make versus female victims across major US cities. Most estimates are around or below 1, and many confidence intervals cross 1, indicating that in most cities male-victim homicides are solved slightly lower rates than female-victim homicides, and there is not significance difference in the odds of solving homicides. Cities like New York, NY, encounter significant lower odds of solving homicides for male victims compared to female victims.

# Problem 2
```{r}
library(p8105.datasets)
data("weather_df")

set.seed(123)

boot_once = function(data) {
  boot_df = data %>% 
    filter(name == "CentralPark_NY") %>% 
    sample_frac(size = 1, replace = TRUE)
    fit = lm(tmax ~ tmin + prcp, data = boot_df)
    r2_hat = broom::glance(fit)$r.squared
    coef_df = broom::tidy(fit)
    beta1 = coef_df$estimate[coef_df$term == "tmin"]
    beta2 = coef_df$estimate[coef_df$term == "prcp"]
    ratio = beta1 / beta2
    tibble(r2 = r2_hat, ratio = ratio)
}

boot_res = map_dfr(1:5000, ~boot_once(weather_df))

boot_res %>%
  pivot_longer(everything()) %>%
  ggplot(aes(value)) +
  geom_density() +
  facet_wrap(~name, scales = "free") +
  labs(title = "Bootstrap distributions of r^2 and Beta1/Beta2") + 
  theme_minimal()
```

The bootstrap distribution of $\hat{r^2}$ is smooth and almost symmetric, centered around 0.915. This means that the fitted model consistently explains a large amount of the variation in `tmax`, and the estimated $\hat{r^2}$ does not change much across different bootstrap samples.
In contrast, the bootstrap distribution of the ratio $\frac{\hat{\beta_1}}{\hat{\beta_2}}$ is extremely skewed. Most values are around zero, and a very long tail exntends to very large numbers. This suggests that ${\hat{\beta_2}}$ is often very close to zero in bootstrap samples, causing the ratio to explode and making the statistic highly unstable. The coefficient ratio is strongly influenced by sampling noise.


```{r}
ci_r2 = quantile(boot_res$r2, probs = c(0.025, 0.975))
ci_ratio = quantile(boot_res$ratio, probs = c(0.025, 0.975))

tibble(parameter = c("r^2", "beta1 / beta2"),
       lower = c(ci_r2[1], ci_ratio[1]),
       upper = c(ci_r2[2], ci_ratio[2])) %>% 
  knitr::kable(digits = 4)
```

# Problem 3
```{r}
birthweight = read_csv("data/birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate(babysex = factor(babysex, levels = c(1, 2),
                          labels = c("male", "female")),
         frace = factor(frace, levels = c(1, 2, 3, 4, 8, 9),
                        labels = c("White", "Black", "Asian", 
                                   "Puerto Rican", "Other", "Unknown")),
         mrace = factor(mrace, levels = c(1, 2, 3, 4, 8),
                        labels = c("White", "Black", "Asian", "Puerto Rican", "Other")),
         malform = factor(malform, levels = c(0, 1), 
                          labels = c("absent", "present")))

birthweight %>% summarize(across(everything(), ~ sum(is.na(.))))

bw_model = lm(bwt ~ babysex + bhead + blength + gaweeks + wtgain + mheight + momage + mrace + ppbmi, data = birthweight)
summary(bw_model)
```

I proposed a regression model that includes the predictors `babysex`, `bhead`, `blength`, `gaweeks`, `wtgain`, `mheight`, `momage`, `mrace`, and `ppbmi`. This model incorporates both infant characteristics and maternal factors that are biologically relevant to birthweight. Baby’s head circumference, height, gestational age, and sex are the primary and most direct indicators of a baby's size and development at birth. The height, weight gain, and BMI of the mother reflect nutritional status and overall growth environment for the baby, which may also influence the baby's birthweight. In addition, I also added the mother's age and race in the model to capture any potential demographic differences that could have an effect on birth outcomes.

```{r}
library(modelr)
birthweight %>% 
  add_predictions(bw_model) %>%
  add_residuals(bw_model) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5, size = 0.8) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(x = "Fitted birthweight (grams)", y = "Residuals",
       title = "Residuals vs fitted values") +
  theme_minimal()
```

The residuals appear roughly centered around zero. A few extreme residuals indicate potential outliers that the model does not capture well.

```{r}
# Cross validation
set.seed(123)

cv_df = crossv_mc(birthweight, n = 100, test = 0.2) %>%
  mutate(train = map(train, as_tibble),
         test  = map(test,  as_tibble))

cv_models = cv_df %>% 
  mutate(model_prop = map(train, 
                        ~ lm(bwt ~ babysex + bhead + blength + gaweeks + 
                               wtgain + mheight + momage + mrace + ppbmi,
                             data = .x)),
         model_main = map(train, ~ lm(bwt ~ blength + gaweeks, data = .x)),
         model_int = map(train, ~ lm(bwt ~ bhead * blength * babysex, data = .x)))

rmse_calc = function(model, test_df) {
  preds = predict(model, newdata = test_df)
  sqrt(mean((test_df$bwt - preds)^2))
}

cv_results = cv_models %>%
  mutate(rmse_new   = map2_dbl(model_prop, test, rmse_calc),
         rmse_main  = map2_dbl(model_main,  test, rmse_calc),
         rmse_int  = map2_dbl(model_int,  test, rmse_calc))

cv_results %>%
  summarize(mean_rmse_new  = mean(rmse_new),
            mean_rmse_main = mean(rmse_main),
            mean_rmse_int = mean(rmse_int))
```

Comparing the mean RMSE of three models, the proposed model with predictors `babysex`, `bhead`, `blength`, `gaweeks`, `wtgain`, `mheight`, `momage`, `mrace`, and `ppbmi` has the lowest average RMSE (275.5092), indicating the best predictive performance among the three. The interaction model performs slightly worse (average RMSE = 288.3849), while the main-effects model has the highest average RMSE (330.7093).

